---
title: "swimteam_proj"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(readr)
library(dplyr)
library(lubridate)
library(anytime)
library(ggplot2)
library(reshape2)
```

```{r}
swim <- read_csv("CleanedSwim.csv")
swim$Timestamp <- anydate(swim$Timestamp)
head(swim)
```

```{r}
swim$week <- format(swim$Timestamp, "%U")

swim$week <- as.numeric(swim$week)

swim <- swim %>% 
  mutate(week = week - 41)

swim <- swim %>% 
  mutate(week = if_else(week < 0, week + 49, week))
```

#mess of a graph
```{r}
#swim_oneperson <- swim %>% 
#  filter(Email == 6)
#swim_oneperson <- swim_oneperson[c(5, 7, 26)]
#
#d <- melt(swim_oneperson, id.vars="week")
#
#ggplot(data=d, aes(week, value)) +
# geom_point(aes(color = variable))

```

#PARTICIPATION FUNCTIONS
```{r}
#Shoulderarm
swim$shoulder_arm_participate[swim$shoulder_arm_participate == 'Full participation without shoulder/arm problems'] <- 0
swim$shoulder_arm_participate[swim$shoulder_arm_participate == 'Full participation, but with shoulder/arm problems'] <- 8
swim$shoulder_arm_participate[swim$shoulder_arm_participate == 'Reduced participation due to shoulder/arm problems'] <- 17
swim$shoulder_arm_participate[swim$shoulder_arm_participate == 'Cannot participate due to shoulder/arm problems'] <- 25
```

```{r}
#backneck
swim$backneck_participate[swim$backneck_participate == 'Full participation without neck/mid back problems'] <- 0
swim$backneck_participate[swim$backneck_participate == 'Full participation, but with neck/mid back problems'] <- 8
swim$backneck_participate[swim$backneck_participate == 'Reduced participation due to neck/mid back problems'] <- 17
swim$backneck_participate[swim$backneck_participate == 'Cannot participate due to neck/mid back problems'] <- 25
```

```{r}
#lowback
swim$lowback_participate[swim$lowback_participate == 'Full participation without low back problems'] <- 0
swim$lowback_participate[swim$lowback_participate == 'Full participation, but with low back problems'] <- 8
swim$lowback_participate[swim$lowback_participate == 'Reduced participation due to low back problems'] <- 17
swim$lowback_participate[swim$lowback_participate == 'Cannot participate due to low back problems'] <- 25
```

```{r}
#hip
swim$hip_participate[swim$hip_participate == 'Full participation without hip/pelvis problems'] <- 0
swim$hip_participate[swim$hip_participate == 'Full participation, but with hip/pelvis problems'] <- 8
swim$hip_participate[swim$hip_participate == 'Reduced participation due to hip/pelvis problems'] <- 17
swim$hip_participate[swim$hip_participate == 'Cannot participate due to hip/pelvis problems'] <- 25
```

```{r}
swim$shoulder_arm_participate <- as.numeric(swim$shoulder_arm_participate)
swim$backneck_participate <- as.numeric(swim$backneck_participate)
swim$lowback_participate <- as.numeric(swim$lowback_participate)
swim$hip_participate <- as.numeric(swim$hip_participate)
```



#REDUCTION FUNCTION
```{r}
helperFunction_red <- function(x){
  swim <- swim %>%
      mutate({{x}} := if_else(swim[[x]] == "No reduction", 0, if_else(swim[[x]] == 'To a minor extent', 6, if_else(swim[,x] == 'To a moderate extent', 12,if_else(swim[,x] == "To a major extent", 18, 25)))))
    return(swim)
    
}
```

```{r}
swim <- helperFunction_red('shoulder_arm_reduction')
swim <- helperFunction_red('backneck_reduction')
swim <- helperFunction_red('lowback_reduction')
swim <- helperFunction_red('hip_reduction')
```

#PERFORMANCE EFFORANT FUNCTION
```{r}
helperFunction_pe <- function(x){
  swim <- swim %>%
      mutate({{x}} := if_else(swim[,x] == "No effect", 0, if_else(swim[,x] == 'To a minor extent', 6, if_else(swim[,x] == 'To a moderate extent', 12,if_else(swim[,x] == "To a major extent", 18, 25)))))
    return(swim)
    
}
```

```{r}
swim <- helperFunction_pe('shoulder_arm_performeff')
swim <- helperFunction_pe('backneck_performeff')
swim <- helperFunction_pe('lowback_performeff')
swim <- helperFunction_pe('hip_performeff')
```

#PAIN FUNCTION
```{r}
helperFunction_pain <- function(x){
  swim <- swim %>%
      mutate({{x}} := if_else(swim[[x]] == "No pain", 0, if_else(swim[[x]] == 'Mild pain', 8, if_else(swim[,x] == 'Moderate pain', 17, 25))))
    return(swim)
    
}
```

```{r}
swim <- helperFunction_pain('shoulder_arm_pain')
swim <- helperFunction_pain('backneck_pain')
swim <- helperFunction_pain('lowback_pain')
swim <- helperFunction_pain('hip_pain')
```



#COMBINE PAIN INDEX
#CREATE A DIFFERNCE OF EXPECTED SWIM YARDS

```{r}
swim <- swim %>%
  mutate(total_pain = shoulder_arm_pain + backneck_pain + lowback_pain + hip_pain,
         target_actual_dif = actual_swim_yards - target_swim_yards) %>% 
  relocate(target_actual_dif, .after = actual_swim_yards)
```

```{r}
#write.csv(swim, 'swim_for_shiny.csv')
```

```{r}
swim_individual <- swim %>% 
    filter(Email == 24)

target <- paste("Target", swim_individual$Email[1], ".png", sep = "")
target
```


#TARGET VS ACTUAL plots
```{r}
swim_individual <- swim %>%  
    filter(Email == 13)

```

```{r}
ggplot(swim_individual, aes(x= week, y= target_actual_dif)) + 
  geom_hline(yintercept=0, linetype="dashed", color = "red") + 
  geom_path() + 
  geom_point(size = 3, color = 'dark blue') + 
  scale_y_continuous(limits = c(-30000, 30000)) + 
  scale_x_continuous(breaks = seq(1,15,by=1), limits = c(0,16)) +
  theme(panel.grid.minor.x = element_blank()) +
  theme_bw() +
  labs(y = "actual yards swam above (or below) target ",
       title = "Target vs Actual",
       subtitle = "more than 0 = over target, less than 0 = under target")

  
#target <- paste("Target", swim_individual$Email[1], ".png", sep = "")
#ggsave(target) 
```


#pain index plots
```{r}
indpain <- swim[c(2,13, 17, 21, 25, 26, 27)]

indpain <- pivot_longer(indpain, cols= shoulder_arm_pain:hip_pain, names_to = 'paintype', values_to = 'painval')

```

```{r}
#write.csv(indpain, 'swim_for_shiny_pain.csv')
```

```{r}

pain_graph <- ggplot(indpain, aes(x= week, y= painval, label = sum(painval ))) + 
  geom_col(aes(fill = paintype), width = 0.85, position = 'stack') +
  scale_x_continuous(breaks = seq(1, 16, by=2), limits = c(0,16))
  #geom_text(aes(label = painval), size = 3, hjust = 0.5, vjust = 3, position = "stack") 

pain_graph

#pain <- paste("Pain", swim_individual$Email[1], ".png", sep = "")
#ggsave(pain)  
```


```{r}
proportion <- ggplot(indpain, aes(x=week,y=painval, group=paintype, colour=paintype)) + 
  geom_line(size = 1.2, alpha = 0.5, position=position_jitter(w=0.4, h=0.3))
proportion
```



#week checking
```{r}
check <- swim %>% 
  select(Timestamp, week, Email) %>% 
  arrange(Email)
check
```

#Modeling...
```{r}
model1 <- lm(shoulder_arm_pain~shoulder_arm_participate+shoulder_arm_reduction+shoulder_arm_performeff, data = swim)
summary(model1)
```

#work on linear models for pain compared to Q1-3
#see what pain levels do to the other questions level

#models of effort
#models on target vs actual

#update graph and add total of q1-3


#APRIL 13

#graphs on one pane
#change values of no reports (signify no report)
#simplify
#create info on weeks of no response

#trend for whole team